Report on laboratory assignments 1 and 2 for EDA282, \\ Parallel Computer Organization and Design

### 2014-10-30 lab/report/include/lab1.tex
revision_1

In the producer consumer benchmark, the E state is beneficial if the writing thread accesses the data element for the first time or when it has been fetched from higher level caches. Then, the thread will transition the data element from I to M through E, when performing an increment. This avoids the S state, which give an increased performance as no bus command has to be issued. However, the probability that the writing thread accesses the data first is low, $\frac{1}{nthreads}$, which is shown in \reffig{fig:lab1bars} where MSI and MESI seems to perform equally. As the working set size increases, the benefit of the E state increases due to more frequent reads from globally invalid cache lines. This can however not be seen in the simulation results, due to fluctuations.

In the no sharing benchmark the first read access to every element end in state S for MSI and state E for MESI, respectively. The following write transitions the element to state M for both the CCPs, but MSI has to issue a BusUpdate whereas MESI's transition is costless. The BusUpdate broadcasts an invalidate signal which gives MSI a slight reduction in performance compared to MESI. In the small working set the elements will stay in state M to the end of execution, but in working sets larger than a private L1 cache the element has to be fetched from higher levels. This makes MSI issue an invalidate broadcast for each increment and the accumulated overhead is easily spotted in \reffig{fig:lab1bars}.

@@

The producer consumer benchmark benefit from the S state, and therefor produces MESI better results than MESI-MG. In the benchmark, one thread update an element whereas the other threads read it and by migrating the element from one private cache to the other is less optimal than letting the reading threads share the element and read it in parallel. This overhead will be a smaller part of the total cycles as the working set increases, as the misses increase in the private caches and also in the LLC, which is costly in terms of cycles. This trend is shown in \reffig{fig:lab1bars}.

>>>>

\textcolor{red}{In the producer consumer benchmark, the E state is beneficial if the writing thread accesses the data element before any other thread, during one iteration when the data element has been fetched from higher level caches. This happens on the first iteration and on every iteration when the element does not reside in cache due to an earlier conviction. If the writing thread succeed with accessing the data element first, it will transition the data element from state I to M through E, when performing an increment. This avoids the S state, which give an increased performance as no bus command has to be issued. However, the probability that the writing thread accesses the data first is low, $\frac{1}{nthreads}$, which is shown in \reffig{fig:lab1bars} where MSI and MESI seems to perform equally. As the working set size increases, the benefit of the E state magnifies due to that the data elements more frequently will be convicted from cache before they are accessed again. The benefit can, however, not be seen in the simulation results, due to fluctuations.}

In the no sharing benchmark the first read access to every element end in state S for MSI and state E for MESI, respectively. The following write transitions the element to state M for both the CCPs, but MSI has to issue a BusUpdate whereas MESI's transition is costless. The BusUpdate broadcasts an invalidate signal which gives MSI a slight reduction in performance compared to MESI. In the small working set the elements will stay in state M to the end of execution, but in working sets larger than a private L1 cache the element has to be fetched from higher levels. This makes MSI issue an invalidate broadcast for each increment and the accumulated overhead is easily spotted in \reffig{fig:lab1bars}. \textcolor{red}{Table \ref{tbl:lab1:task3} express the overhead in more detail when the work set size increase, where MSI's write misses and invalidation commands increase as much as MESI's write hits.}

@@

The producer consumer benchmark benefit from the S state, and therefor produces MESI better results than MESI-MG. In the benchmark, one thread update an element whereas the other threads read it and migrating the element from one private cache to the other is less optimal than letting the reading threads share the element and read it in parallel. This overhead becomes a smaller part of the total cycles as the working set increases, because the misses increase in the private caches and also in the LLC, which is costly in terms of cycles. This trend is shown in \reffig{fig:lab1bars}.
